---
output:
  html_document:
    toc: false
    theme: cerulean
    highlight: zenburn
    keep_md: true
---

```{r knitr_setup, echo=FALSE}
opts_chunk$set(cache=TRUE, echo=FALSE, eval=TRUE, tidy=FALSE, message=FALSE, warning=FALSE, results="asis", comment=NA)
read_chunk("../../code/cc4lite.R")
```

```{r setup}
```

```{r functions}
```

##
##
### Spatial scale: precipitation

Now repeat the previous comparisons, this time for precipitation.
Due to the heteroskadasticity in the precipitation variable, the patterns of change in mean and variability between scales are similar.
Only the floating range bars are shown here.

##
#### Mean precipitation: 2-km resolution

```{r plot08a_cru_precip_exclusive_2km}
```

##
#### Mean precipitation: 10-minute resolution

```{r plot08b_cru_precip_exclusive_10min}
```

Comparisons between data sets for Fairbanks, Alaska suggest there is little difference between resolutions.
For wetter locations such as Juneau, Alaska, heteroskedastic properties allow for more difference in gross precipitation amounts,
though other characteristics such as the seasonal cycle appear preserved across resolutions.

##
#### Precipitation range: 2-km resolution

```{r plot09a_cru_precip_exclusive_2km}
```

##
#### Precipitation range: 10-minute resolution

```{r plot09b_cru_precip_exclusive_10min}
```

<style>iframe.rChart{ width: 100%; height: 500px;}</style>

In summary, the effect of spatial scale is stronger on precipitation than temperature.
At least for the communities considered, temperature differences are relatively negligible, as long as the interest remains in broad trends over long periods of time.
Precipitation should be compared across scales with even greater caution.
