---
output:
  html_document:
    toc: false
    theme: cosmo
    highlight: zenburn
    keep_md: true
---

```{r knitr_setup, echo=FALSE}
opts_chunk$set(cache=FALSE, eval=FALSE, tidy=TRUE, message=FALSE, warning=FALSE)
read_chunk("../../code/agg_city_data.R")
```

Most of the QAQC documentation and code can be explored under that project.
Here I will discuss only the most relevant end stage **R** code used to curate the data for direct use in the Community Charts application,
specifically the script, `agg_city_data.R`.

### Code flow
Here is a diagram for a fuller context regarding how the Community Charts application grows out of the SNAP data QAQC project.

```{r code_sankey, echo=F, eval=T}
```

```{r code_sankey_embed, echo=F, eval=T, comment=NA, results="asis", tidy=F}
```

## R code

### Setup
ADD_TEXT_HERE: EXAMPLE
Setup consists of loading required **R** packages and additional files, preparing any command line arguments for use, and defining functions and other **R** objects.

