---
output:
  html_document:
    toc: false
    theme: cerulean
    highlight: zenburn
    keep_md: true
---

```{r knitr_setup, echo=FALSE}
opts_chunk$set(cache=FALSE, eval=FALSE, tidy=TRUE, message=FALSE, warning=FALSE)
read_chunk("../../code/agg_city_data.R")
```

##
##

Most of the QAQC documentation and code can be explored under that project.
Here I will discuss only the most relevant end stage **R** code used to curate the data for direct use in the Community Charts application,
specifically the script, `agg_city_data.R`.

### R code

General setup includes specifying paths, loading **R** packages, setting the number of processing cores for `parallel`, defining decades of interest, and listing files.

```{r setup}
```

The function used for loading and manipulating the data into curated form.
The `%>%` (pipe operator) is available from the `magrittr` package, which is loaded when `dplyr` is loaded.

```{r function}
```

The data are processed in parallel.
The lines which are commented out refer to some tests of compute times for different methods of combining multiple data frames.

```{r process}
```

A variety of useful **R** workspace files are saved.
Some prior workspaces are also loaded at specific stages below to be added to the current workspace before it is saved again.

```{r save}
```
